P.berghei pir analysis

### 2021-09-14 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
change pir proportion Pberg graph by facetting for sub-fam and including genome composition

  tpm_data_subfam <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    filter(Geneid %in% pir_id) %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(clade = pir_clade$clade[match(Geneid, 
                                       pir_clade$Geneid)]
         ) %>% 
  group_by(clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
    mutate(stack_group = clade) %>% 
    group_by(stage, 
             stack_group,
             clade) %>% 
    dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(clade),
          total_tpm
)

tpm_data_subfam <- mutate(tpm_data_subfam,
                   order = row_number())

  # #Apply ordering
  # tpm_data_subfam$stage = factor(tpm_data_subfam$stage, 
  #                         levels = ordering(tpm_data_subfam$stage,
  #                                           order_vector))

  #Calculate the range for the y value to make it look nicer
  stacked_max <- tpm_data_subfam %>% 
    ungroup %>% 
    group_by(stage) %>% 
    dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
    .$max_tpm %>% 
    max

  max_tpm <- c(
    seq(
      0,
      mround(
        stacked_max,
        500),
      length.out = 11)
  )
```

```{r combined_tpm_stacked_plot}

@@

y <- ggplot(data = tpm_data_subfam, 
            aes(
              x = stage,
              y = total_tpm, 
              fill = clade,
              group = order
            )) +
  geom_bar(
    # position = position_stack(),
    position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1) +
  scale_fill_manual(
    values = clade_col_num
  ) +
  scale_color_manual('black') + 
  labs(fill = stack_label) +
  scale_y_continuous(breaks = seq(0,1,0.1)) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('Proportion of '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~stage, 
             scales = "free_x",
             space = "free_x") +
  scale_x_discrete(expand=expand_scale(add=1)) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x = element_blank(),
    # axis.line.x = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=8, 
                                face = 'bold', 
                                angle = 90),
    strip.background = element_rect(
      color="black", 
      size=0.1, 
      linetype="solid"
    ),
    panel.spacing.x = unit(0, "null")
  )

>>>>

tpm_data_subfam <- combined_data %>% 
  rownames_to_column(var = 'Geneid') %>% 
  filter(Geneid %in% pir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'stage', 
       value.name = 'tpm') %>% 
  mutate(clade = pir_clade$clade[match(Geneid, 
                                       pir_clade$Geneid)],
         SorL = pir_clade$SorL[match(Geneid, 
                                       pir_clade$Geneid)]
  ) %>% 
  group_by(SorL, clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
  group_by(stage,
           SorL,
           clade) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>%
  group_by(stage) %>% 
  mutate(prop_tpm = total_tpm/sum(total_tpm)) %>% 
  arrange(
    desc(clade),
    total_tpm
  )

clade_composition <- pir_clade %>% 
  group_by(SorL, clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
  summarise(prop_tpm = n_distinct(Geneid)/nrow(.)) %>% 
  mutate(stage = "Genome") %>% 
  melt(id.vars = c('stage', 'SorL', 'clade'), 
       value.name = 'prop_tpm') %>% 
  select(-variable)

tpm_data_subfam <- bind_rows(tpm_data_subfam, clade_composition) %>% 
  mutate(stage = factor(stage, 
                        levels = ordering(unique(c(as.character(tpm_data_subfam$stage), 
                                                   clade_composition$stage)), 
                                          order_vector = c(r_order_vector, 
                                                           'composition'))))

#Calculate the range for the y value to make it look nicer
stacked_max <- tpm_data_subfam %>% 
  ungroup %>% 
  group_by(stage) %>% 
  dplyr::summarise(max_tpm = sum(prop_tpm)) %>% 
  .$max_tpm %>% 
  max

max_tpm <- c(
  seq(
    0,
    mround(
      stacked_max,
      500),
    length.out = 11)
)

@@

y <- ggplot(data = tpm_data_subfam, 
            aes(
              x = stage,
              y = prop_tpm, 
              fill = clade
            )) +
  geom_bar(
    position = position_stack(),
    # position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1) +
  scale_fill_manual(
    values = clade_col_num
  ) +
  scale_color_manual('black') + 
  labs(fill = stack_label) +
  scale_y_continuous(breaks = seq(0,1,0.2)) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('Proportion of '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(cols = vars(stage),
             rows = vars(SorL),
             scales = "free_x",
             space = "free_x") +
  scale_x_discrete(expand=expand_scale(add=1)) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x = element_blank(),
    # axis.line.x = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=8, 
                                face = 'bold', 
                                angle = 90),
    strip.background = element_rect(
      color="black", 
      size=0.1, 
      linetype="solid"
    ),
    panel.spacing.x = unit(0, "null")
  )
